---
title: "Phylogenetic cleaning"
output: pdf_document
bibliography: ../manuscript/hiDrosophoBioGeo.bib
---

```{r setup, include=FALSE}
library(knitr)

opts_chunk$set(echo = FALSE)

load('keepTrack.RData')
```

We are using the time-calibrated phylogeny from @magnacca2015, specifically the Newick file provided as Supplementary data 1.  Below are a few questions about matching the phylogeny to the spatial data

## Definition of the picture wing clade

What do we want to do with *Drosophila primaeva* and *D. adventitia*? According to @magnacca2015 these are both outside the picture wing clade.  However, we have spatial data for *D. primaeva* (the sister of the picture wings); we do not have spatial data for *D. adventitia* (sister to the clade comprised of picture wings plus *D. primaeva*).

## Non-matching names

Some names from the spatial data are not found in the tree, and conversely some names in the tree are not found in the spatial data.  Currently, there are `r sum(keepTrack$miss[, 1] != '')` species in the spatial data that do not have a match in the phylogeny.  These species account for `r round(100 * keepTrack$nlost, 2)`% of the spatial records. Can we further match any of these taxa (e.g. by synnonomy)?

Below is a table (in alphabetical order) of names from one source that don't match in the other.

```{r}
kable(keepTrack$miss)
```

## References
